---
title: "Lesson 4 - Reading Data and first data structures"
output: html_document
---

# Read CSV data

Read data using read.csv function

Why cats1? We will look at different versions
for education purposes.

```{r}
cats1 <- read.csv("data/feline-data-v1.csv")
```

We want to view it

```{r}
cats1
```

```{r}
str(cats1)
```

# Extract column

```{r}
cats1$weight
```

I can compute things, e.g. convert to gram.

```{r}
cats1$weight * 1000
```

Save this result in the data frame
```{r}
cats1$weight_g <- cats1$weight * 1000
cats1
```

What will this give:

```{r}
cats1$weight + cats1$coat
```

This is an error because non-matching data types

# How to find out data types



The function for this is called typeof

```{r}
typeof(cats1$weight)
```

```{r}
typeof(cats1$coat)
```

Try out other values we have seen before:

```{r}
typeof( 3 > 2 )
```

Counts have a special type

```{r}
nrow(cats1)
typeof(nrow(cats1))
```

# Import an updated cats dataset

```{r}
cats2 <- read.csv("data/feline-data-v2.csv")
cats2
```

Calculate the weight in grams

```{r}
cats2$weight_g <- cats2$weight * 1000
```

```{r}
str(cats2)
```

# Our second data structure: Vector

What is a vector

```{r}
vec <- c(1,2,3,4)
vec
cats1$weight # similar
```

```{r}
str(vec)
```

We can calculate with vectors

```{r}
vec * 2
```

Create character vectors

```{r}
cat_names <- c("Boo", "Naila", "Romeo")
cat_names
```

Combine

```{r}
weights <- c(2.1, 5, 3.2, "2.3 or 2.4")
weights
```

compeare this to 

```{r}
str(cats2)
```


# Resolve it: Clean up your data

```{r}
cats3 <- read.csv("data/feline-data-v3.csv")
cats3
```


This data set has only **atomic values**! No "..or..."

 Never glue together values!
 
 * "2.3. or 3.4"
 * "2020_control"
 * "04/01/2022"  -> split this in day, month year, that are atomic.
