---
title: "Lesson 06  - Subsetting data"
output: html_document
---

# Subsetting data in vectors

```{r}
x <- c(5.4, 6.2, 7.1, 4.8, 7.5)
x
```

Single values: Use the bracket notation
```{r}
2
x[2]
```

Use a vector (2) to subset another vector (x)

Extend this to multiple values: 

- Subset elements 1 and 3

```{r}
x[ c(1, 3) ]
```

- Subset the values 1 to 4

```{r}
1:4
x[1:4]
```

Indices are error-prone and not well to imagine

-> Use names

```{r}
#var = 2 # equals to var <- 2
pizza_price <-
  c(margherita = 4.50,
    tonno = 5.60,
    funghi = 5.20, 
    capricciosa = 7.20)
```

Get Funghi price: Use "character" index instead of a
numeric index -> less room for error!

```{r}
pizza_price["funghi"]
```

## Omitting values

All values except the second:

```{r}
x
x[-2]
```


Challenge:

Combine:

Omit from x the 3rd to the 5th value

(Remember: " 3 to 5" == 3:5)
```{r}
-(3:5) # change the operator precedence!
x [  -(3:5)  ]
```

Remove and save

```{r}
x <- x[-2]
x
```

# Subset by logical


Create a logical vector

```{r}
logi <- c(T, F, F, T, F) # omit this in scripts (F <- 3 ?!)
logi <- c(TRUE, TRUE, FALSE, TRUE, FALSE) #< this is the way

```

Subset with it

```{r}
x
x[logi] # logical too long
x[c(TRUE, FALSE)] # logical too short -> recycles (T, F, T, F)
x[c(TRUE, TRUE, FALSE, FALSE)]
```

How to generate logical vectors?

```{r}
x < 5
```

-> I can filter! FOr example: Filter out element of
x that are smaller than 5

```{r}
x[ x < 5 ]
```

# More complex filtering: Combine multiple logical statements:

 * Condition 1 is TRUE AND condition 2 is TRUE:  ->  &
 * Condition 1 is TRUE OR  condition 2 is TRUE:  ->  |
 * Invert Condition 1:   !    (e.g.  !(3 > 4) -> TRUE     )
Example: TRUE & TRUE   -> TRUE
         TRUE & FALSE  -> FALSE
         
**Challenge: Filter elements from x that are smaller than 7
    AND bigger than 5
    
```{r}
x <- c()
```
    
```{r}
# TIP: Create the logical vector first by combining
# questions about x (e.g. x > 5)
sel <- ( x < 3 ) | ( x > 4 )
x[sel]
# expected result: 5.4
```
Bonus question1: Achieve the same thing with | 

```{r}
x[ ( ... )  | ( .... ) ]
```

Bonus question2:  Achieve the same thing with | and & and
  WITHOUT flipping < and >
  
```{r}
x[  ...( x > 5 ) | ...(x < 7)  ]
```


```{r}

```







