---
title: "Control Flow"
author: "C. Beleites"
date: "21 1 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Controlling Flow: if- / if-else-statement

```
if (condition) {
  do_if_true()
}
```

```{r}
x <- 8

if (x >= 10) {
  print("x is greater than or equal to 10")
}
```

this is how R (de)constructs the statement part in curly braces

from "inside out"
```{r}
# innermost
"x is greater than or equal to 10"
# if we don't say what function to use, R will assume show()

# we do have a function here that is applied to this value:
print("x is greater than or equal to 10")

# the outermost part of the statement are the curly brackets:
{
  print("x is greater than or equal to 10")
}
```


now deconstruct the condition part analogously:
```{r}
x

x >= 10
```



if we want to execute some statements if the condition is *not* met:

```{r}

# reformulate the condition
if (x < 10) {
  print("x is not greater than or equal to 10")
}

# careful: does ! x >= 10 mean !(x >= 10) or (!x) >= 10
if (!(x >= 10)) {
  print("x is not greater than or equal to 10")
}
```

Suppose we want to do something if the condition is met, and something
else if it is not met:

```{r}
if (x >= 10) {
  print("x is greater than or equal to 10")
} else {
  print("x is less than 10")
}

```

Pitfall: variables can store logical values, which are then evaluated as condition

```{r}
x  <-  4 == 3
if (x) {
  "4 equals 3"
} else {
  "4 does not equal 3"
}
```

## x being a vector (with multiple elements) in the condition

```{r}
x <- 1 : 10

if (x >= 10) {
  print("x is greater than or equal to 10")
} else {
  print("x is less than 10")
}

```

3 ways of what may need to be done

1. to run the conditional statement on each of the vector elements
2. to execute the `TRUE` branch if *all* elements of the logical vector are `TRUE`
3. to execute the `TRUE` branch if *any* elements of the logical vector is `TRUE`

```{r}
any(x >= 10)
all(x >= 10)
```

```{r}
if (any (x >= 10)) {
  print("some value(s) in x is/are greater than or equal to 10")
} else {
  print("all values in x are less than 10")
}
```

