---
title: "Functions"
author: "C. Beleites"
date: "21 1 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#' square function
#' 
#' This function takes a parameter x and squares it.
#'
#' @param x numeric, to be squared
#' @param print logical, says whether to make additional printout of the square
#'
#' @return numeric, the square of x
#'
#' @examples
#' square(3)
#' square(10 : 20)
square <- function(x, print = TRUE) {
  x <- x^2
  
  if (print) {
    cat("The square of x is", x, "\n")
  }
  
  ## you can also explain what parts of your function do:
  return(x)
}

square(2)
```

```{r}
x <- 1 : 10

sqrt(square(-3))

absolute_value <- function(x) {
  x <- square(x, print = FALSE)
  x <- sqrt(x)
  
  return(x)
}

absolute_value(-3)

x

y <- -1 : 3
absolute_value(y) # call by value (as opposed to call-by-reference)
y

```


compare: the x inside `absolute_value()` is -3 in our call, even though in the global workspace x is a vector of numbers from 1 to 10.

## Possibilities

- use `source()` to source out function definitions into .R script files
- functions are just a particular type of variables
- and can be handed over as parameter to other functions
  `f(s)` -> `use_on_all_vector_elements(vector, f)` doing f(vector[i]) along the whole vector.
  That is the `*apply` family of functions.

