---
title: "Lesson 1 - Introduction"
output: html_document
---

Create new chunk with Ctrl + Alt + i

```{r}
1 + 100
```


More calculations

```{r}
2 /4
```

Parenthesis change the operator precedence

```{r}
(2 + 3) * 2 
```


Scientific notation

```{r}
3 * 10000000  # 3 * 10^7 = 3e07
```

# Mathematical functions

```{r}
sin(pi)
```


Logarithm

```{r}
log10(1000)
```


# Compare

Are 2 things equal?
```{r}
2 == 4
```

Missing values

```{r}
 2 == NA
```

Other comparison (Evaluate whole chunk with Ctrl + Shift + Return)

```{r}
3 > 2
2 >= 2
5 <= 7
3 != 4


```


Combining TRUE and FALSE

```{r}
(3 < 5) & (1 != 2)
```


# Variables

Give results a name

```{r}
addition_part   <- 2 + 3
addition_part * 2
```

But using another factor gives another result

```{r}
addition_part * 3
```


What names can you use?

Must not

* Contain spaces
* start with an underscore
* have math operators
* start with a number

use Alt -  to insert <-


```{r}
min_height <- 2
# _age <- 2  
.mass <- 3 # hidden!
MaxLength <- 3
# min-length <- 3
# 2widths <- 2
celsius2kelvin <- 3
```



# Managing your environment

delete variables

Best way: => Delete environment and
 Recompute everything!

Haphazard way:

```{r}
rm(min_height)
m <- MaxLength
(m*2) - (m*1) / (m^2)
rm(m)
```





